ms_figures

### 2018-11-19 man/figures/bai_cronos.Rmd
update

```{r echo=FALSE, results='asis', eval=show_text}
cat("
# Lista de figuras

- [] Figure 1. Location map

    - [] Ver https://github.com/ajpelu/qpyr_dendro/blob/master/out/fig/geomap/mapa_situacion.pdf

- [] Figure 2. Plot de anomalias
- [] Figure 3. Resilience metrics
- [] Figure 4. BAI chronologies

- [] Figure 5. Residual chronologies --> Supplementary

")
```

**Figure 1**. Distribution of *Quercus pyrenaica* forests in the Iberian Peninsula (a) and in Sierra Nevada mountain range (b). Different colours indicate oak population cluster's identified in Sierra Nevada [@PerezLuque2015]. For each population, a grid with the MODIS pixels is shown (see material and methods). Detailed location of the dendroecological sampling sites: northern (San Juan, SJ) (c), and southern ones (Cáñar: CA-Low and CA-High) (d). Colour orthophotography of 2009 from Regional Ministry of the Environment [@IECA2009].

![](../out/fig/geomap/mapa_situacion.jpg){height=600px}

##### 
**Figure 2.** EVI standardized anomaly during the period 2000-2016 for northern and southern populations. Error bars show standard error.

```{r, eval=FALSE}
iv <- read.csv(here("data/evi", "iv_composite.csv"), header = TRUE, sep = ',')

drought_years <- c(2005, 2012)

evi_profile_dat_ref <- iv %>% 
  filter(!year %in% drought_years) %>% 
  group_by(composite) %>% 
  dplyr::summarise(mean=mean(evi, na.rm=TRUE),
            sd = sd(evi, na.rm=TRUE),
            se = sd/sqrt(length(evi))) %>% 
  mutate(composite_dates = 
           plyr::mapvalues(composite,
                           c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23),
                           c('01-01','01-17','02-02','02-18','03-06','03-22','04-07','04-23',
                             '05-09','05-25','06-10','06-26','07-12','07-28','08-13','08-29',
                             '09-14','09-30','10-16','11-01','11-17','12-03','12-19'))) %>%
  mutate(cd = as.Date(composite_dates, format = '%m-%d')) %>% 
  mutate(period = 'reference')

evi_profile_dat_2005 <- iv %>% 
  filter(year == 2005) %>% 
  group_by(composite) %>% 
  dplyr::summarise(mean=mean(evi, na.rm=TRUE),
            sd = sd(evi, na.rm=TRUE),
            se = sd/sqrt(length(evi))) %>% 
  mutate(composite_dates = 
           plyr::mapvalues(composite,
                           c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23),
                           c('01-01','01-17','02-02','02-18','03-06','03-22','04-07','04-23',
                             '05-09','05-25','06-10','06-26','07-12','07-28','08-13','08-29',
                             '09-14','09-30','10-16','11-01','11-17','12-03','12-19'))) %>%
  mutate(cd = as.Date(composite_dates, format = '%m-%d')) %>% 
  mutate(period = '2005')

evi_profile_dat_2012 <- iv %>% 
  filter(year == 2012) %>% 
  group_by(composite) %>% 
  dplyr::summarise(mean=mean(evi, na.rm=TRUE),
            sd = sd(evi, na.rm=TRUE),
            se = sd/sqrt(length(evi))) %>% 
  mutate(composite_dates = 
           plyr::mapvalues(composite,
                           c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23),
                           c('01-01','01-17','02-02','02-18','03-06','03-22','04-07','04-23',
                             '05-09','05-25','06-10','06-26','07-12','07-28','08-13','08-29',
                             '09-14','09-30','10-16','11-01','11-17','12-03','12-19'))) %>%
  mutate(cd = as.Date(composite_dates, format = '%m-%d')) %>% 
  mutate(period = '2012')

# Join the three dataframes
evi_profile_compara_dat <- rbind(evi_profile_dat_ref, evi_profile_dat_2005, evi_profile_dat_2012)

# colref <- '#455883'
# colref2 <- '#10253F'
# col2005 <- '#00BA38'
# col2012 <- '#F8766D'

col2012 <- '#0700fe'
col2005 <- '#19e00b'
colref <- '#7b7b7b'

profile_compara <- ggplot(evi_profile_compara_dat, aes(cd, y=mean, color=period)) + 
  geom_errorbar(aes(ymin = mean - 2*se, ymax= mean + 2*se), width=3, size=.5) + 
  #geom_errorbar(aes(ymin = mean - sd, ymax= mean + sd), width=4, colour='black') + 
  geom_line(size=.9) + 
  geom_point(size=1.5, fill='white', shape=21) +
  scale_x_date(labels = function(x) format(x, "%b"),
               breaks = date_breaks('month')) + 
  ylab('EVI') + xlab('Date') + 
  theme_bw() +
  theme_classic() + 
  theme(panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        legend.position = c(0.1,.9),
        plot.title = element_text(hjust = -0.05, size = 16, 
                                    margin = margin(t=-5, b=1))) +
  scale_color_manual(values=c(col2005, col2012, colref))

# profile_compara 
```

```{r}
# From explore_anomalies.Rmd 
# Read data 
anomalias_evimean <- read.csv(here("data/anomalies", "anomalias_evimean.csv"),header = TRUE, sep = ',')

anomalias_evimean <- anomalias_evimean %>% 
  mutate(
    clu_pop = as.factor(case_when(
      pop == 1 ~ "Camarate",
      pop %in% c(2,3,4,5) ~ 'Northern slope',
      pop %in% c(6,7,8) ~ 'Southern slope',
      pop == 9 ~ 'out')),
    clu_pop2 = as.factor(case_when(
      pop %in% c(1,2,3,4,5) ~ 'Northern slope',
      pop %in% c(6,7,8) ~ 'Southern slope',
      pop == 9 ~ 'out'))) %>% 
  filter(clu_pop != 'out')

# colours 
color_neg <- '#a63603'
color_pos <- '#006d2c'

# Standardized Anomalies 
avg_sa_clu <- anomalias_evimean %>% 
  group_by(clu_pop2, y) %>% 
  dplyr::summarise(mean=mean(sa, na.rm=T),
            sd = sd(sa, na.rm=T),
            se = sd/sqrt(length(sa))) %>% 
 mutate(signo = ifelse(mean >= 0, 'pos', 'neg'))

myylab <- 'EVI Standardized Anomaly'

plot_sa_clu  <- avg_sa_clu %>%  
  ggplot(aes(x=y, y=mean, fill=signo)) + 
  geom_bar(stat = "identity") + 
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se, colour=signo), width=.5) +
  facet_wrap(~clu_pop2, nrow = 2) + 
  scale_fill_manual(values = c("pos" = color_pos, "neg" =  color_neg)) +
  scale_color_manual(values = c("pos" = color_pos, "neg" =  color_neg)) +
  ylab(myylab) + xlab('') +
  theme_bw() +
  theme(text = element_text(size=16),
        legend.position = "none",
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        strip.background = element_rect(fill = "white"),
        plot.title = element_text(hjust = -0.05, size = 16, 
                                    margin = margin(t=-5, b=1)))

# plot_sa_clu
```

```{r, fig.height=9} 
# grid.arrange(profile_compara, plot_sa_clu)
```

```{r, fig.height=7, eval=FALSE}
a_profile_compara <- profile_compara + labs(title = "a")
b_plot_sa_clu <- plot_sa_clu + labs(title = "b")
a_profile_compara + b_plot_sa_clu + plot_layout(ncol=1)
```

```{r, fig.height=5}
plot_sa_clu 
```

##### 
**Figure 3.** Response *Q. pyrenaica* forests to drought in terms of resistance, recovery and resilience of greenness (EVI; left-plots) and tree radial growth (BAI; right-plots) for the years 2005 and 2012. For EVI we compared northern populations (*black fill circle*) with southern ones (*blue empty circle*). For BAI we compared northern population (San Juan, SJ; *black triangle*) with southerns populations: Cáñar-High (CA-High; *blue empty squares*) and Cáñar-Low (CA-Low; *blue fill squares*). Different letters above error bars indicate significant *post hoc* differences between groups (see material and methods).

```{r, eval=FALSE, echo=FALSE}
mhuber <- read.csv(here("/out/anovas_resilience/huber_evi", "robust_mhuber.csv"), header = TRUE)
mhuberbai <- read.csv(here("/out/anovas_resilience/huber_bai", "robust_mhuber.csv"), header = TRUE)

# NO plot rrs 
mhuber_filter <- mhuber %>% filter(var != 'rrs') %>% mutate(disturb_year = as.factor(disturb_year))
mhuberbai_filter <- mhuberbai %>% filter(var != 'rrs') %>% mutate(disturb_year = as.factor(disturb_year))

pd <- position_dodge(.2)
line_lwd <- .8

aes_disturb <- aes(x = disturb_year, y = M.Huber, 
                   color = site, group = site, 
                   fill = site, linetype=site,
                   shape = site, 
                   ymin = lower.ci, ymax = upper.ci)
aes_site <- aes(x = site, y = M.Huber, 
                color = disturb_year, group=disturb_year, 
                fill=disturb_year, linetype=disturb_year,
                shape = disturb_year, 
                ymin = lower.ci, ymax = upper.ci)

theme_res <- list(
  geom_errorbar(width=.2, position=pd, lwd=line_lwd), 
  geom_line(position = pd, lwd=line_lwd),
  geom_point(size = 2.5, position = pd), 
  facet_wrap(
    ~var_sorted, nrow = 1, scales="free_y",
    labeller = as_labeller(c(
      "0_rt" = "Resistance",
      "1_rc" = "Recovery",
      "2_rs" = "Resilience"))),
  geom_text_repel(aes(label=Letter), 
                  nudge_x = - 0.5,
                  nudge_y = 0.01,
                  direction="both", 
                  size = 3.5, show.legend = FALSE,
                  segment.size = 0.5, 
                  segment.alpha = 0.5),
  theme_bw(), 
  theme(
    panel.grid.minor = element_blank(),
    strip.background = element_rect(
      colour = "black",
      fill = "white"),
    plot.title = element_text(hjust = -0.05, size = 16, margin = margin(t=-5, b=1))),  
  ylab("Response value"), 
  xlab("")
)

theme_evi <-  list(
  scale_color_manual(values = c("black", "blue")),
  scale_shape_manual(values=c(19, 22)),
  scale_fill_manual(values = c("black", "blue")),
  scale_linetype_manual(values = c("solid", "solid"))
)

theme_bai <- list(
  scale_color_manual(values=c('black','blue','blue')), 
  scale_shape_manual(values=c(19, 22, 15)),
  scale_fill_manual(values=c('black','white','blue')),  
  scale_linetype_manual(values=c("solid", "solid", 'solid'))
)
```

```{r, eval=FALSE, echo=FALSE}
# EVI_drought 
res_evi_drought <-
  ggplot(mhuber_filter, aes_disturb) + 
  theme_res + theme_evi + theme(legend.position='top') + 
  ggtitle("a) EVI")

# BAI_drought 
res_bai_drought <-
  ggplot(mhuberbai_filter, aes_disturb) + 
  theme_res + theme_bai + theme(legend.position='bottom') + 
  xlab('Drought event') + 
  ggtitle("b) BAI")

```

```{r, eval=FALSE, fig.height=7}
res_evi_drought  + res_bai_drought + plot_layout(ncol=1)
```

```{r, fig.height=9}
mhuber <- read.csv(here("/out/anovas_resilience/huber_evi", "robust_mhuber.csv"), header = TRUE)
mhuberbai <- read.csv(here("/out/anovas_resilience/huber_bai", "robust_mhuber.csv"), header = TRUE)

# NO plot rrs 
mhuber_filter <- mhuber %>% filter(var != 'rrs') %>% mutate(disturb_year = as.factor(disturb_year))
mhuberbai_filter <- mhuberbai %>% filter(var != 'rrs') %>% mutate(disturb_year = as.factor(disturb_year))

aux <- mhuber_filter %>% mutate(vari = '0_EVI')
aux2 <- mhuberbai_filter %>% mutate(vari = '1_BAI')
m <- bind_rows(aux2, aux)

pd <- position_dodge(width = 0.6)

aes_disturb <- aes(x = disturb_year, y = M.Huber, 
                   color = site_sorted, group = site_sorted,  
                   fill = site_sorted, linetype=site_sorted,
                   shape = site_sorted, 
                   ymin = lower.ci, ymax = upper.ci)

variables  <- c("0_EVI" = "EVI", "1_BAI" = "BAI")

# orden <- c("N", "S", "SJ", "CA-High","CA-Low")
item_site <- c("0_N" = "N", "1_S" = "S", "2_SJ" = "SJ", "3_CA-High" = "CA-High", "4_CA-Low" = "CA-Low")

tl <- guide_legend(title = "Site")

themeR <- list(
  geom_errorbar(width=.1, position = pd),
  geom_point(position = pd),
  theme_bw(),
  facet_wrap(~vari, labeller = labeller(vari = variables), scales="free"),
  xlab(''), 
  theme(
    panel.grid.minor = element_blank(),
    strip.background = element_rect(
      colour = "black",
      fill = "white"),
    strip.text = element_text(face="bold", size = 14),
    axis.title.y = element_text(face="bold", size = 14),
    plot.title = element_text(hjust = -0.05, size = 16, margin = margin(t=-5, b=1))),
  scale_shape_manual(labels = item_site, guide = guide_legend(title = "Site"),
    values=c("0_N" = 19, "1_S" = 21, "3_CA-High" = 22, "4_CA-Low" = 15, "2_SJ" = 17)), 
  scale_color_manual(labels = item_site, guide = tl,
    values=c("0_N" = 'black', "1_S" = "blue", "3_CA-High" = 'blue', "4_CA-Low" = 'blue', "2_SJ" = 'black')),
  scale_fill_manual(labels = item_site, guide = tl,
    values=c("0_N" = 'black', "1_S" = "white", "3_CA-High" = 'white', "4_CA-Low" = 'blue', "2_SJ" = 'black')),
  scale_linetype_manual(labels = item_site, guide = tl,
                        values=c("solid","solid","solid","solid", "solid")),
  scale_size_manual(labels = item_site, guide = tl,
    values=c("0_N" = "2", "1_S" = "2", "3_CA-High" = 2.5, "4_CA-Low" = 2.5, "2_SJ" = 2.5)),

  geom_text(aes(label=Letter, y=postext), 
            position=pd, 
            show.legend = FALSE, size = 3)
  # , position_jitterdodge(dodge.width = 0.6, jitter.width = .2)
  # geom_text_repel( 
  #                 nudge_x = - 0.5,
  #                 nudge_y = 0.01,
  #                 direction="both", 
  #                 size = 3.5, show.legend = FALSE,
  #                 segment.size = 0.5, 
  #                 segment.alpha = 0.5) 
)

m <- m %>% mutate(postext = case_when(
  vari == "0_EVI" ~
    case_when(var_sorted == "0_rt" ~ 0.97,
              var_sorted == "1_rc" ~ 1.19,
              var_sorted == "2_rs" ~ 1.02),
  vari == "1_BAI" ~
     case_when(var_sorted == "0_rt" ~ 1.1,
              var_sorted == "1_rc" ~ 1.7,
              var_sorted == "2_rs" ~ 1.2)))

m <- m %>% mutate(site = as.factor(
  case_when(
  site == "caH" ~ "CA-High",
  site == "caL" ~ "CA-Low",
  TRUE ~ site
)))

m <- m %>% mutate(site_sorted = as.factor(
  case_when(
  site == "N" ~ "0_N",
  site == "S" ~ "1_S",
  site == "SJ" ~ "2_SJ",
  site == "CA-High" ~ "3_CA-High",
  site == "CA-Low" ~ "4_CA-Low"
)))

rt <- m %>% filter(var == "rt") %>% 
  ggplot(aes_disturb) + themeR + 
  ylab("Resistance") 
# + geom_text(aes(label=Letter), position=position_jitterdodge(jitter.width = .2, dodge.width = 0.6), show.legend = FALSE)

rc <- m %>% filter(var == "rc") %>% 
  ggplot(aes_disturb) + themeR + 
  ylab("Recovery")
# + geom_text(aes(label=Letter, y = 1.65), position=pd, show.legend = FALSE)

rs <- m %>% filter(var == "rs") %>% 
  ggplot(aes_disturb) + themeR + 
  ylab("Resilience") + xlab('Drought event') 
# + geom_text(aes(label=Letter, y = 1.2), position=pd, show.legend = FALSE) 
```

```{r, fig.height=6}
grid_arrange_shared_legend(rt, rc, rs,ncol = 1, nrow = 3, position='top') 
```

```{r, eval=FALSE}
### Figure 4

theme_de <- list(
  scale_color_manual(values=c('black','red')), 
  scale_shape_manual(values=c(19, 15)),
  scale_fill_manual(values=c('black','red')),  
  scale_linetype_manual(values=c("solid", "solid")),
  labs(color = 'Drougth event', fill='Drougth event', linetype='Drougth event', shape='Drougth event')

)

# EVI_site
res_evi_site <-
  ggplot(mhuber_filter, aes_site) + 
  theme_res + theme_de + theme(legend.position='top') + 
  ggtitle("a) EVI")

# BAI_site
res_bai_site <-
  ggplot(mhuberbai_filter, aes_site) + 
  theme_res + theme_de + theme(legend.position='bottom') + 
  xlab('Site') + 
  ggtitle("b) BAI") 
```

```{r, eval=FALSE}
res_evi_site  + res_bai_site + plot_layout(ncol=1)

**Figure 4.** Response of *Q. pyrenaica* forests to 2005 (*black*) and 2012 (*red*) drought events in terms of resistance, recovery and resilience of greenness (EVI; a) and tree radial growth (BAI; b) by site. Different letters above bars indicate significant post hoc differences between groups (see material and methods). 
```

##### 
```{r, warning=FALSE}
cro_sites <- read.csv(here("data/cronos_medias", "cronos_sites.csv"), header = TRUE, sep = ',')

@@

**Figure 4.** Basal Area Increment (BAI) chronologies of *Q. pyrenaica* for northern population (SJ; *green*) and southern ones: low-elevation (CA-Low; *pink*) and high-elevation (CA-High, *purple*) sites. Shading areas correspond to standard error of the mean. Number of series are displayed in the upper plot. We only show years replicated with # series > 5. Linear trend since 1975 is shown for southern high-elevation site (CA-High) ($r_2$ = `r r2caH`).

@@

##### 
**Figure 5.** Residual tree-ring chronologies obtained for the *Q. pyrenaica* sites. Dashed red lines indicate the start of the reliable period (EPS > 0.85). Dotted black lines showing the three of most recent severe drought years (1995, 2005 and 2012).

```{r}
cro_rwi <- read.csv(here("data/cronos_medias", "cronos_sites_rwi.csv"), header = TRUE, sep = ',')
year_cutoffs <- read.csv(here("data/cronos_medias", "cronos_sites_rwi_year_cutoffs.csv"), header = TRUE, sep = ',')
year_sampledepth <- read.csv(here("data/cronos_medias", "cronos_sites_rwi_year_sampledepth.csv"), header = TRUE, sep = ',')

# peores sequias 
dy <- read.csv(file=here::here('data/sequias', 'severe_spei12.csv'), header=TRUE)
dy <- dy %>% arrange(desc(d_duration))
# solamente aquellos mayor de 2 meses 
dys <- dy %>% filter(d_duration > 2) %>% filter(maxyear > 1940 ) %>% as.data.frame() 
dys <- dys[,'maxyear'] 
```

```{r}
# top10  
top10pos50 <- cro_rwi %>% filter(year > 1950) %>% arrange_(~desc(res)) %>% group_by_(~site) %>%  slice(1:10)
top10neg50 <- cro_rwi %>% filter(year > 1950) %>% arrange_(~res) %>% group_by_(~site) %>%  slice(1:10)
top10pos <- cro_rwi %>% arrange_(~desc(res)) %>% group_by_(~site) %>%  slice(1:10)
top10neg <- cro_rwi %>% arrange_(~res) %>% group_by_(~site) %>%  slice(1:10)
```

```{r}
label_site <- c("0_sj" = "SJ",
                "1_caH" = "CA-High",
                "2_caL" = "CA-Low")

cro_rwi <- cro_rwi %>% 
  mutate(siteF = case_when(
    site_sorted == "0_sj" ~ "Northern (SJ)",
    site_sorted == "1_caH" ~ "Southern (Ca-High)",
    site_sorted == "2_caL" ~ "Southern (Ca-Low)"))

# 
plot_rwi_crono <- ggplot(cro_rwi, aes(x=year, y=res)) +
  theme_bw() + 
  theme(panel.grid = element_blank(), 
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face="bold", size=12), 
        # strip.text = element_blank(), 
        axis.text = element_text(face="bold", size=8), 
        axis.title = element_text(face="bold", size=12)) + 
  geom_hline(yintercept = 1, colour = "grey") + 
  geom_line(size=.8) + scale_y_continuous(breaks = seq(0.4,1.8, 0.2)) +
  facet_wrap(~site_sorted, scales = "free_x", ncol = 1,
             labeller = as_labeller(label_site)) + 
  ylab("Ring width index") + xlab("Year") + 
  geom_vline(aes(xintercept = year), data = year_cutoffs,  linetype = "dashed", colour='red') + 
  # geom_vline(aes(xintercept = year), data = year_sampledepth,  linetype = "dashed", colour='blue') +
  geom_vline(xintercept = c(1995, 2005, 2012), linetype = "dotted") + 
  # geom_vline(xintercept = dys, linetype = "dotted") + 
  scale_x_continuous(limits = c(1819, 2016), 
                     breaks= seq(1820, 2020, by=20))

plot_rwi_crono  
```

##### 
**Figure 6.** Correlation coefficients obtained by relating tree-ring residual chronologies (RWI) of *Q. pyrenaica* and monthly climatic data: precipitation (a), SPEI (b), maximun (c) and minimun (d) temperatures. *green* bars: northern site (SJ); *light blue* bars: low-elevation southern site (CA-Low); and *dark blue* bars: high-elevation shouthern site (CA-High). Asteriks indicate significant ($P < 0.05$) correlation coefficients.

```{r}
rtmax <- read.csv(here("data/rwi_climate", "rtmax_selected.csv"), header = TRUE, sep = ',')
rtmin <- read.csv(here("data/rwi_climate", "rtmin_selected.csv"), header = TRUE, sep = ',')
rsequia <- read.csv(here("data/rwi_climate", "rsequia_selected.csv"), header = TRUE, sep = ',')
rprec <- read.csv(here("data/rwi_climate", "rprec_selected.csv"), header = TRUE, sep = ',')

rtmax <- rtmax %>% mutate(v_name = 'T max')
rtmin <- rtmin %>% mutate(v_name = 'T min')
rsequia <- rsequia %>% mutate(v_name ='SPEI')
rprec <- rprec %>% mutate(v_name = 'Prec')

df <- rbind(rtmax, rtmin, rsequia, rprec)

# cutre solucion para el problema de PREC 
exdf <- data.frame(
  id = c(16:19),
  varname = rep("prec", 4), 
  month = c('dec...FEB', 'MAR...MAY', 'JUN...AUG', 'SEP...NOV'),
  coef = rep(0, 4), 
  significant = rep('FALSE', 4),
  ci_lower = rep(0, 4),
  ci_upper = rep(0, 4), 
  sig = rep("", 4), 
  month_name = c("Winter", "Spring", "Summer", "Autumn"), 
  v_name = rep("Prec", 4))

exdf <- rbind(exdf, exdf, exdf)
exdf$X <- c(46:57)
exdf$site <- c(rep("SJ", 4), rep("caL", 4), rep("caH", 4))

ddf <- rbind(df, exdf)

label_var <- c("Prec" = "(a) Prec", "SPEI" = "(b) SPEI", 
               "T max" = "(c) T max", "T min" = "(d) T min")

label_site <- c("SJ" = "SJ",
                "caH" = "CA-High",
                "caL" = "CA-Low")

clima_ring <- ddf %>% ggplot(aes(x=id, y=coef, group = site)) + 
  geom_hline(yintercept = 0, colour="gray") +
  geom_bar(aes(fill=site), stat="identity", position = "dodge") +
  facet_wrap(~v_name, labeller = as_labeller(label_var)) + 
  geom_text(aes(label = sig, x = id, y = ifelse(coef > 0, (coef + 0.02), (coef - 0.05))), 
            position = position_dodge(.9), size = 6) + 
  theme_bw() + 
  theme(axis.title.x = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.3),
        legend.position = "bottom",
        strip.background = element_rect(fill=NA, colour=NA),
        strip.text.x = element_text(size = 10, face = "bold")) + 
  scale_x_continuous(breaks = unique(ddf$id), labels = unique(ddf$month_name)) +
  scale_y_continuous(limits = c(-0.4, 0.8), 
                     breaks = c(-0.4, -0.2, 0, 0.2, 0.4, 0.6,0.8)) + 
  scale_fill_manual(values = c("#B2DF8A","#1F78B4","#A6CEE3"), labels = label_site) + 
  ylab('Correlation')

clima_ring 
```

#####
**Figure 7.** Comparison of median growth change ($GC$) following @Nowacki1997 for *Q. pyrenaica* sites. Dashed black lines indicate a threshold of 50 % of GC (see material and methods).

```{r}
gc_todos <- read.csv(here("data/disturbance", "disturbance_gc.csv"), header = TRUE, sep = ',')

# customize plot (corrected scale)
pos_neg_log <- scales::trans_new('signed_log', 
                                 transform = function(x) sign(x)*log(abs(x)),
                                 inverse = function(x) sign(x)*exp(abs(x)))

custom_ylab <- '% GC'
micolor <- c("#838B8B")

# solution for arrange
# http://felixfan.github.io/stacking-plots-same-x/

custom_corrected_gc <- list(
  geom_bar(stat='identity', fill=micolor, colour=micolor),
  theme_bw(), 
  geom_hline(yintercept=c(50, -50), linetype=2, colour='black'),
  theme(panel.grid.minor = element_blank(), 
        strip.background = element_rect(colour='black',fill='white'))
  )

c <- gc_todos %>% 
  filter(site != 'gc_sj') %>% 
  ggplot(aes(x=year, y=gc_md)) + 
  geom_hline(yintercept = 0, colour = 'black', lwd = 0.3) + 
  geom_errorbar(aes(ymin = gc_md - gc_md_se, ymax = gc_md + gc_md_se), colour='gray', width=.3, size=.2) + 
  # scale_y_continuous(trans = pos_neg_log) +
  facet_wrap(~site_sorted, nrow = 2,
             labeller=as_labeller(c("1_caH" = "CA-High", "2_caL" = "CA-Low"))) + 
    ylab(custom_ylab) + custom_corrected_gc

d <- gc_todos %>% 
  filter(site == 'gc_sj') %>% 
  ggplot(aes(x=year, y=gc_md)) + 
  xlim(layer_scales(c)$x$range$range[1],layer_scales(c)$x$range$range[2]) +
  geom_hline(yintercept = 1, colour = 'gray', lwd = 0.3) + 
  geom_errorbar(aes(ymin = gc_md - gc_md_se, ymax = gc_md + gc_md_se), colour=micolor, width=.3, size=.2) +
  facet_wrap(~site_sorted, nrow = 2,
             labeller=as_labeller(c("0_sj" = "SJ"))) + 
  custom_corrected_gc + ylab('') + 
  scale_y_continuous(trans = pos_neg_log, 
                     breaks=c(-1000, -500, -100, 0, 100, 500)) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank())

plot_medians_corregido <- d + c + plot_layout(ncol=1, heights = c(1, 2))
plot_medians_corregido

```

##### 
**Figure 8**. Percentage of *Q. pyrenaica* trees affected by GC > 50 % by site. *Black* line shows number of trees (rigth-axis).

```{r}
gc_treesMed <- read.csv(here("data/disturbance", "disturbance_trees_gc_treshold_median.csv"), header = TRUE, sep = ',')
gc_treesMed50 <- read.csv(here("data/disturbance", "disturbance_trees_gc_treshold_median50.csv"), header = TRUE, sep = ',')

colorpositivo <- c("#104E8B")
colornegativo <- c("#CD3333")

pgc_median50 <- gc_treesMed50 %>% ggplot() +
  geom_bar(aes(x=year, y=per_pos), fill=colorpositivo, colour=colorpositivo, stat = "identity", position = "dodge") + 
  geom_bar(aes(x=year, y=(-1)*per_neg), fill=colornegativo, colour=colornegativo, stat = "identity", position = "dodge") +
  facet_wrap(~site_sorted, nrow = 3, labeller = as_labeller(label_site)) + 
  scale_y_continuous(sec.axis = sec_axis(~ ., name = '# trees', breaks = c(0,15,30))) +
  geom_line(aes(x=year, y=total_tree), colour = 'black') + 
  theme_bw() + ylab('% trees') +
  geom_hline(yintercept=c(50, -50), linetype=2, alpha=.4) +
  theme(panel.grid.minor = element_blank(), 
        strip.background = element_rect(colour='black', 
                                        fill='white')) 
pgc_median50

```
##### 
**Appendix S1.** Temporal evolution of cumulative precipitation (hydrological year) during the period 1950-2017. Points represent mean and errorbars standard error. *Black* line indicates mean for all period. *Red* lines represent -1 and -2 standard deviation (*dotted* and *dashed* lines respectively). *Blue* lines represent +1 and +2 standard deviation (*dotted* and *dashed* lines respectively). Years with average values below -1SD are labelled. Data from 28 meteorological stations distributed around Sierra Nevada area (from National Spanish Meteorological Services, AEMET). ***Inset plot***: cumulative precipitation during the hydrological years 2004-2005 (*blue line*) and 2011-2012 (*red line*). The boxplot representing the average from 1950-2015 period. Data from meteorological station Granada, Base Aérea.

```{r, fig.height=5}
ltdf <- read.table(file=here("/data_raw/meteo", "long_term_data_rediam_data.txt"), header=T, sep=';', dec=',')

source(here("/script/R/", "hydro_year.R"))

ltdf <- ltdf %>% 
  dplyr::select(cod = INDICATIVO, VARIABLE, FECHA, VALOR) %>% 
  dplyr::mutate(fecha = as.Date(FECHA, format='%d/%m/%Y'),
         month = lubridate::month(fecha),
         hmonth = ifelse(month <= 8, month + 4, month -8), 
         hyear = hydro_year(fecha),
         hyear_f = paste0(hyear-1,'-', hyear))

# We used only the 5514 station (the most long term)
# Filter by station and prec 
# Filter from 1950 
st5514 <- ltdf %>% 
  filter(VARIABLE == 'PM1') %>% 
  filter(cod == '5514') %>% 
  filter(hyear >= 1950) %>% 
  dplyr::select(fecha, month, hmonth, hyear, hyear_f, pre=VALOR)

st5514_pro <- st5514 %>% 
  group_by(hyear_f, hmonth) %>% 
  dplyr::summarise(pre = sum(pre)) %>%
  dplyr::mutate(csum = cumsum(pre)) %>% as.data.frame()

df <- st5514_pro  %>% 
  filter(!(hyear_f %in% c('2004-2005', '2011-2012')))

cp <- ggplot(df, aes(x=as.factor(hmonth), y=csum)) + 
  geom_boxplot(aes(x=as.factor(hmonth))) +  
  geom_smooth(aes(group=1), se=TRUE, level=0.95, colour='black') + 
  geom_line(data=subset(st5514_pro, hyear_f == '2004-2005'),
            aes(x=as.factor(hmonth), y=csum, group=1), colour='blue') +
  geom_line(data=subset(st5514_pro, hyear_f == '2011-2012'),
            aes(x=as.factor(hmonth), y=csum, group=1),
            colour='red') +
  theme_bw() + 
  theme(strip.background = element_rect(fill = NA, color = "black"),
        axis.text.x = element_text(angle = 90, hjust = 1, size=7),
        axis.text.y = element_text(size=7),
        axis.title.y = element_text(size = 7)) + 
  ylab('Cumulative Precipitation (mm)')  + xlab('') + 
  # ggtitle('Granada Base Aerea, 1950-2015') + 
  scale_x_discrete(labels=c("1"="Sep", "2"="Oct", "3"="Nov", "4"="Dec", "5"="Jan", "6"="Feb",
                            "7"="Mar", "8"="Apr", "9"="May", "10"="Jun", "11"="Jul", "12"="Aug")) +
  annotate("text", label = "2004-2005", x = 2, y = 500, size = 2, colour = "blue") + 
  annotate("text", label = "2011-2012", x = 2, y = 550, size = 2, colour = "red") + 
  annotate("text", label = "1950-2015", x = 3, y = 600, size = 2, colour = "black") +
  theme(panel.grid = element_blank())
```

```{r}
prec_files <- list.files(here::here('data_raw/meteo/prec_subclima/'), pattern="*.csv")

dfpre <- c()

for (i in prec_files){

  auxdf <- read.csv(here::here('data_raw/meteo/prec_subclima', i), header = TRUE, sep=";")
  auxdf$code <- gsub(x=i, pattern = ".csv", replacement = "")

  dfpre <- rbind(dfpre, auxdf)
}

dfpre$date <- as.Date(dfpre$date, format = "%d/%m/%Y")

dfpre <- dfpre %>% 
  dplyr::mutate(fecha = as.Date(date, format='%d/%m/%Y'),
         month = lubridate::month(fecha),
         hmonth = ifelse(month <= 8, month + 4, month -8), 
         hyear = hydro_year(fecha),
         hyear_f = paste0(hyear-1,'-', hyear))

aux <- dfpre %>% 
  group_by(hyear_f, code) %>% 
  dplyr::mutate(csum = cumsum(prec)) %>% 
  filter(hmonth == 12) %>% 
  dplyr::select(-month, -prec, -date, -fecha) %>% as.data.frame()

presn <- aux %>% group_by(hyear) %>% 
     dplyr::summarise(
     csum_mean = mean(csum, na.rm = TRUE),
     csum_sd = sd(csum, na.rm = TRUE), 
     csum_se = csum_sd/sqrt(length(csum_sd)), 
     n = n())

meandf <- mean(presn$csum_mean)
sddf <- sd(presn$csum_mean)

>>>>

##### 
**Figure 4.** Basal Area Increment (BAI) chronologies of *Q. pyrenaica* for northern population (SJ; *green*) and southern ones: low-elevation (CA-Low; *pink*) and high-elevation (CA-High, *purple*) sites. Shading areas correspond to standard error of the mean. Number of series are displayed in the upper plot. We only show years replicated with # series > 5. Linear trend since 1975 is shown for southern high-elevation site (CA-High) .AREGLAR ESTO
##### 
```{r, warning=FALSE}
cro_sites <- read.csv(here("data/cronos_medias", "cronos_sites.csv"), header = TRUE, sep = ',')

@@

@@

### 2018-11-19 man/ms_figures.Rmd
last update
update

```{r, eval=FALSE}
res_evi_site  + res_bai_site + plot_layout(ncol=1)

**Figure 4.** Response of *Q. pyrenaica* forests to 2005 (*black*) and 2012 (*red*) drought events in terms of resistance, recovery and resilience of greenness (EVI; a) and tree radial growth (BAI; b) by site. Different letters above bars indicate significant post hoc differences between groups (see material and methods). 
```

>>>>

```{r, eval=FALSE}
res_evi_site  + res_bai_site + plot_layout(ncol=1)
```
